---
title: "COVID19 Vaccination Rate Mini Project"
author: 'Ari_Fon (PID: A15390446)'
date: "3/3/2022"
output: 
  pdf_document:
    latex_engine: xelatex
---

Here we downloaded the data from file called “Statewide COVID-19 Vaccines Administered by ZIP Code” using website https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code

#First step is to import/read vaccination data

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q1. What column details the total number of people fully vaccinated?

persons_fully_vaccinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

2021-01-05

>Q4. What is the latest date in this dataset?

2022-03-01

```{r}
vax$as_of_date[nrow(vax)]
```

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

18338

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 



```{r}
round((18338/107604)*100, 2)

```


>Q8. [Optional]: Why might this data be missing?

Not everyone gets them from the CDC, such as people who get them somewhere else such as the VA or elsewhere. 


#Working with Dates

```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

```{r}
age <-today()-ymd("2000-01-20")
age
```

```{r}
time_length(age, "year")
```


>Q9. How many days have passed since the last update of the dataset? 

time difference of 2 days
```{r}
today()-vax$as_of_date[nrow(vax)]
```


>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

61 unique dates in the dataset

```{r}
udates <- unique(vax$as_of_date)
length(udates)
```

#Working with Zipcodes

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```


```{r}
reverse_zipcode(c('92037', "92109") )
```



```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode(vax$zip_code_tabulation_area )
head(zipdata)
```




##Focus on the San Diego area

Let’s now focus in on the San Diego County area by restricting ourselves first to vax$county == "San Diego" entries. We have two main choices on how to do this. The first using base R the second using the dplyr package:

```{r}
library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```


>Q11. How many distinct zip codes are listed for San Diego County? 

107 distinct zip codes listed for SD county
```{r}
uzip <- unique(sd$zip_code_tabulation_area)
length(uzip)
```

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

92154

```{r}
sdmax12 <- which.max(sd$age12_plus_population)
print(sdmax12)
sd$zip_code_tabulation_area[sdmax12]
```

Using dplyr select all San Diego “county” entries on “as_of_date” “2022-02-22” and use this for the following questions.


```{r}
sd$as_of_date[nrow(sd)]
```
Let's do this with the most recent date in the data-set (2022-03-01).

>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-03-01”? 

```{r}
#filter to the day
sd.latest <- filter(sd, as_of_date == "2022-03-01")
round(mean(sd.latest$percent_of_population_fully_vaccinated, na.rm = TRUE),2)
```

```{r}
summary(sd.latest$percent_of_population_fully_vaccinated)
```

71% overall average

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-03-01”?


```{r}
hist(sd.latest$percent_of_population_fully_vaccinated, breaks =30)
```


```{r}
library(ggplot2)

ggplot(sd.latest) +aes(percent_of_population_fully_vaccinated)+ geom_histogram()
```


## Focus on UCSD/La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
baseplot <- ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x= "Date", y="Percent Vaccinated", title= "Vaccination rate for La Jolla 92037" )
baseplot
```



#Comparing to similar sized areas

>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

Add a line showing the average vaccination rate for all zip code areas with a population jyst as large as 92037

```{r}
baseplot + labs(title= "Vaccination rate for CA 92037 (UCSD)")
```
Add a line showing the average vaccination rate for all zip code areas with a population jyst as large as 92037

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")

head(vax.36)
```

```{r}
ave.36 <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
ave.36

baseplot + labs(title= "Vaccination rate for CA 92037 (UCSD)") + geom_hline(yintercept=ave.36, linetype= "dashed", col="red")
                                                                
```


>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)

```


>Q18. Using ggplot generate a histogram of this data.

```{r}
library(ggplot2)
ggplot(vax.36)+ aes(percent_of_population_fully_vaccinated) + geom_histogram() +xlim(c(0,1)) + labs(x= "Percent Vaccinated", y= "Count")
```


>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

Both values are below the mean we found above.
0.754>0.724
0.754>0.552

>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)

ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color= "blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccianted",
       title="Vaccination rates across California",
       subtitle="only areas with a population above 36k are shown") +
  geom_hline(yintercept = ave.36, linetype="dashed")
```


>Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

I grduate this quarter so while I would be down to travel, I won't be a student anymore next quarter (sad)





