---
title: 'class10: Genome Informatics'
author: 'Ari_Fon (PID: A15390446)'
date: "2/17/2022"
output: pdf_document
---

#Examine 1000 Genome Data

> Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)?

```{r}
#Read Genotype file from Ensemble

mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
table(mxl$Genotype..forward.strand.)/nrow(mxl)
```

What about a different population? Here we take the British in England and Scotland (GBR)

```{r}
gbr <-read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
table(gbr$Genotype..forward.strand.)/nrow(gbr)
```


##Expression by Genotype Analysis

I want to read my RNA-Seq expression results into R. This file is not a CSV but rather has fields seprated by space. 


```{r}
x <- read.table("rs8067378_ENSG00000172057.6.txt")
head(x)
```

First try at this question. Is the mean expression different based in gneotype?
\
```{r}
x$geno== "G/G"
```

```{r}
summary(x[x$geno== "G/G", "exp"])
```

```{r}
table(x$geno)
```

```{r}
summary(x[x$geno== "A/G", "exp"])
summary(x[x$geno== "A/A", "exp"])
summary(x[x$geno== "G/G", "exp"])
```

##Make a summary overview figure
```{r}
library(ggplot2)

ggplot(x) + aes(geno,exp, fill=geno) + geom_boxplot(notch = TRUE)
```


